---
title: "General Social Survey"
author: "Dominique Repp"
date: "May 11TH 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent") 
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

# Introduction Question 1
	
	<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->
	
	This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.
	
	How are the "voted in 2016" and "voted in 2012" related?  
	$H_0:$ voting in 2016 and 2012 are not related.  
	
	
	
	## Methods
	
	<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
	 use categorical variables, and a chi-squared test to test the hypothesis that voting in 2016 is not independent of voting in 2012
	
	##Results
	
	<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
	
	
	### Descriptive Results
	
	we use a barchart to show the voting trends between 2012 and 2016
	
	```{r}
	barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent" ,xlab= 'voted in 2016') 
	chisqtestGC(~VOTE16+VOTE12, data=mydata4)
	votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
	fisher.test(votetab)
	```
	
	#### Graphical Descriptive Results
	
	<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
	
	```{r}
	barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent" ,xlab= 'voted in 2016') 
	plot(xtabs(~VOTE16+VOTE12, data=mydata4))
	plot(votetab)
	summary(xtabs(~VOTE16+VOTE12, data=mydata4))
	```
	
	
	#### Numerical Descriptive Results
	
	<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!--> 
	```{r}
	xtabs(~VOTE16+VOTE12, data=mydata4)
	```
	
	
	### Inferential Results
	
	<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
	
	```{r}
	chisqtestGC(~VOTE16+VOTE12, data=mydata4)
	votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
	fisher.test(votetab)
	```
	
	$H_a$: if you voted in 2012 it is likely you would vote in 2016
	
	we reject the null hypothesis, we have evidence that voting in 2012 is related to voting in 2016. If you voted in 2012 it is likely you will vote in 2016.
	